\documentclass[a4paper,10pt]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

%Includes "References" in the table of contents
\usepackage[nottoc]{tocbibind}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{setspace}
\onehalfspacing
% \usepackage{cite}
% \usepackage[backend=bibtex]{biblatex}
\usepackage{hyperref}
\bibliographystyle{plain}

\newenvironment{theorem}[1][]{\par\medskip
   \noindent \textbf{Theorem~#1.} \rmfamily}{\medskip}
\newenvironment{assumption}[1][]{\par\medskip
   \noindent \textbf{Assumption~#1.} \rmfamily}{\medskip}

\begin{document}

\title{Ranking: Consistency and Rates}
\author{Dany Haddad}
\date{July 2020}
\maketitle

\section{Introduction}

This is a work in progress document discussing the consistency of
common convex surrogate functions for ranking as well as the
convergence rate of the corresponding empirical risk. First we
summarize the main points of Duchi et al. in this
regard~\cite{duchi-2010-ranking,duchi-2013-ranking} focusing on the
uniform convergence guarantees. Then we discuss some relevant research
directions, in partictular, we are interested in achieving fast
convergence rates for ranking applications.  \footnote{I also point
  out confusion and questions I have in the footnotes.}

\section{Literature Review}

\subsection{The Asymptotics of Ranking Algorithms}

\subsubsection{Overview}

The authors first introduce the ranking problem and describe a
pairwise loss function, equivalent to the pairwise 0-1~ranking
loss. They then define several forms of consistency for ranking
problems and unify them under some conditions
(theorem~1). Surprisingly, common surrogates for the pairwise
0-1~ranking loss function are shown to not be consistent.

The next section discusses how aggregating the pairwise preferences
into lists allows us to define consistent and tractable surrogate loss
functions. The key idea is that the surrogates must be order preserving.
\footnote{The authors describe inconsistency as arising due to a lack of complete preference information. This point is not entirely clear to me since the proof of inconsistency seems to depend on the variability in preference information rather than the lack of information. See Ravikumar et al. \cite{ravikumar-2011-ranking} for a discussion on how a lack of normalization leads to ``non-robustness''.}

Given that our new surrogate loss functions consist of an aggregation
step, it is natural to consider the aggregation in batches. In
particular, the authors describe the loss function as a sum of
U-statistics. However, this is seems mostly to simplify the analysis
rather than to leverage the fact that U-statistics are UMVUEs.
\footnote{The authors mention that they introduce the U-statistics
  based risk functional to be able to analyze the setup without overly
  detailed knowledge of the surrogate loss. But they already assume
  the surrogate is lipschitz continuous and bounded. What more
  information would we need to conduct the analysis without taking the
  U-statistics route?}
The authors then prove a ULLN for the risk
functional they define and show that it is consistent for the
underlying listwise losses.

\subsection{Preliminaries}

The ranking problem can be described generally as ordering a list of
\(m\) items for each query \(q\) in the query space \(Q\) (which we
assume is countable). The is equivalent to providing a real valued
score for each item in the list.

Duchi et al.~\cite{duchi-2013-ranking} consider a general setup where
the item preference information is received in an arbitrary form and
then transformed into a problem specific structure space by an
aggregation function. For example, the preference information $Y_i$
could be a matrix of pairwise preferences and the aggregation function
$s$ is either a simple average, or a transformation into a list of
item scores. $Y_i$ could also be the interactions a user had with a
list of items (clickstream data). We also consider the case where $s$ performs no
aggregation at all. Consider the 0-1 pairwise ranking loss:

\begin{equation}\label{eq:general-loss}
  L(\alpha, A) = \sum_{i<j} A_{ij}1(\alpha_i \leq \alpha_j) + \sum_{i>j} A_{ij}1(\alpha_i < \alpha_j)
\end{equation}

where $A = s(\{Y_i\}_{i=1}^k)$. Note that this is different from the typical 0-1 pairwise ranking loss in that the error \(\alpha_i = \alpha_j\) is penalized only once.
\footnote{They mention that this is a technical detail, but it's not clear to me where in the proofs it is important. Most likely it has to do with the inconsistency results.}

\begin{assumption}[A]
  The distribution of preference information \(\mu_q^k\) converges weakly to some limiting distribution \(mu_q\) for every query \(q\) where \(k\) is the amount of preference aggregation. So, for \(S_k = s(\{Y_i\}_{i=1}^k) \sim \mu_q^k\) as \(k \rightarrow \infty\):

  \begin{equation}
    \mu_q^k \rightarrow^d \mu_q
  \end{equation}

\end{assumption}

\begin{assumption}[B]
  The loss function \(L\) is bounded in \([0,1]\) and continuous.
\end{assumption}

Given the loss function \(L\) we define the loss over a distribution of preference information \(\mu\) as:

\begin{equation}
  \ell(\alpha, \mu) = \int L(\alpha, s) d\mu(s)
\end{equation}

For the case where the preference information \(Y_i\) is a matrix of pairwise preferences and the aggregation function is the average, then this reduces to: \(ell(\alpha, \mu) = L(\alpha, Y_{i,j}^\mu)\) where \(Y_{i,j}^\mu\) is the average preference of item \(i\) over item \(j\) according to the distribution \(\mu\).

We define the ranking risk of our scoring function \(f: Q \rightarrow R^m\) as:

\begin{equation}
  R(f) = \sum_q^Q p_q \ell(f(q), \mu_q)
\end{equation}

Where \(p_q\) is the prior probability of query \(q\). Without loss of generality, let \(p_q\) be decreasing with the query index \(q\).

Similarly, we define the surrogate loss and surrogate risk as:

\begin{align}
  \ell_\varphi(\alpha, \mu) &= \int \varphi(\alpha, s) d\mu(s) \\
  R_\varphi(f) &= \sum_qp_q\ell_\varphi(f(q), \mu_q)
\end{align}

\subsubsection{Inconsistency Results}


\subsubsection{Proving a ULLN for the U-statistics Based Risk Functional}

Define the U-statistics based empirical risk:

\begin{equation}\label{eq:er}
  \hat{R}_{n, \varphi}(f) = \sum_q \frac{\hat{n}_q}{n} \frac{1}{\binom{\hat{n}_q}{k}} \sum_{I_q \in \mathcal{I}_q}\varphi(f(q), s(\{Y_i\}_{I_q}))
\end{equation}

where \(\hat{n}_q\) is the frequency of query \(q\) in a sample of size \(n\) and \(\mathcal{I}_q\) is the set of all combinations of indexes of samples corresponding to query \(q\).

We include conditions I and II as well as assumptions D, E and F from Duchi et al. \cite{duchi-2013-ranking} in the statement of theorem 4.
\begin{theorem}[4]
  For \(\phi \in [0, B_n]\) bounded and \(L_n\)-lipshitz over \(\mathcal{F}_n\),
  if there exists some \(\rho, \beta > 0\) such that:

  \begin{equation}
    p_q = O(q^{-\beta -1})
  \end{equation}
  and the expected surrogate loss at aggegation level \(k\) satisfies for all \(f, k, q, n\):

  \begin{equation}
    \left|\mathbb{E}_q\varphi(f(q), s(\{Y_i\}^k)) - \lim_{k' \rightarrow \infty}\mathbb{E}_q\varphi(f(q), s(\{Y_i\}^{k'}))\right| = O(B_nk^{-\rho})
  \end{equation}

  with \(B_n/k_n^\rho = o(1)\) and \(k_nB_n^{(1+\beta)/\beta}=o(n)\) and in addition we have the following constrain on the covering numbers:

  \begin{equation}
    k_nB_n\sqrt{\log\left( N(\frac{\epsilon}{4L_n}) \right)} = o(\sqrt n)
  \end{equation}

  then we have the uniform convergence guarantee:

  \begin{equation}
    \sup_{f \in \mathcal{F}_n}\left| \hat{R}_{n, \varphi}(f) - R_\varphi(f) \right| \rightarrow^p 0
  \end{equation}


\end{theorem}
The proof for this uniform law of large numbers can be broken down by decomposing the uniform deviation into two terms and controlling each of them:

\begin{equation}\label{eq:uni-dev}
  \sup_{f \in \mathcal{F}_n} \left| \hat{R}_{n, \varphi}(f) - R_\varphi(f) \right| \leq
  \sup_{f \in \mathcal{F}_n} \left| \hat{R}_{n, \varphi}(f) - R_{n, \varphi}(f) \right| +
  \sup_{f \in \mathcal{F}_n} \left| R_{n, \varphi}(f) - R_{\varphi}(f) \right|
\end{equation}

where we denote:\footnote{The factor of \(1/n\) is missing from the paper but is used in the supplementary material.}

\begin{equation}
  R_{\varphi, n} = \sum_q\sum_{l=0}^n l \mathbb{P}_n/n(\hat{n}_q = l) \mathbb{E}(\varphi(f(q), s(\{Y_i\}^{\min(l, k)}))|Q=q)
\end{equation}

The first term (I) on the RHS of inequality \ref{eq:uni-dev} is the deviation of the U-statistic based estimate \(\hat{R}_{n, \varphi}(f)\) from it's expectation \(R_{n, \varphi}(f)\). The second term (II) is the deviation of \(R_{n, \varphi}(f)\) from it's limit \(R_{\varphi}(f)\) as \(n, k \rightarrow \infty\). The second term is controlled by lemma 8 of the supplementary material \cite{duchi-2013-ranking} while the first term is controlled by lemma 10.

Lemma 8 uses the conditions on the the boundedness of \(\phi\) and the aggregation level \(k_n\) as well as a covering number argument to give a deterministic bount on the second term above.

Lemma 10 shows that the first term is \(o_p(1)\) by arguing that the risk functional \(\hat{R}_{n, \varphi}(f)\) satisfies a bounded differences inequality. Applying McDiarmid's inequality as well as a union bound involving the covering number gives the following bound, yielding the claim:

\begin{equation}
  \mathbb{P}\left( \sup_{f \in \mathcal{F}_n} \left| \hat{R}_{n, \varphi}(f) - R_{n, \varphi}(f) \right| \geq \epsilon \right) \leq 2\exp\left( \log N \left( \frac{\epsilon}{4L_n}, n \right) - \frac{n \epsilon^2}{32k^2B_n^2} \right)
\end{equation}

where \(N \left( \frac{\epsilon}{4L_n}, n \right)\) gives the covering number of \(\mathcal{F}_n\).
\footnote{My understanding is that this gives a rate worse than \(O_p(1/\sqrt n)\) since the concentration inequality involves the covering number which is dependent on \(\epsilon\). Unless for the function class under consideration this can be bound uniformly independent of \(epsilon\), this is worse than the typical rate.}

\subsubsection{Potential improvements}
The U-statistics formulation is totally ignored in controlling term
(I), leading to a rate slower than \(O_p(1/\sqrt n)\). For the case of
bipartite ranking, Clemencon et
al.~\cite{clemencon-2008-ranking-rates} derive the typical \(O_p(q/sqrt n)\) rate of
convergence of the empirical risk minimizer to the optimal
optimizer. The approach involves making a low noise assumption (bound
variance of the risk in terms of the expectation) and applying a
Hoeffding decomposition of the U-statistic based risk. The decomposed
terms are controlled by applying the decoupling technique for
degenerate U-statistics.

However, their results are more relevant in a
model selection context since they deal directly with the non-convex
pairwise 0-1 ranking loss. Clemencon et
al.~\cite{clemencon-2008-ranking-rates} also mention (in Remark 7)
that it would be interesting to derive fast rates for convex
surrogates, but that it would be quite technical. They point to
Blanchard et al.~\cite{blanchard-2003-rates} for a similar derivation
for the binary classification case.

Bartlett et al.~\cite{bartlett-2006-risk} derive fast rates for convex
surrogates in the context of binary classification. They also
demonstrate via simulation that their derived rates are
representative. Clemencon et al.~\cite{clemencon-2008-ranking-rates}
show how the results from this work transfer over to the case of
bipartite ranking but do not derive fast rates.

The U-statistics based empirical risk defined in this work
(\ref{eq:er}) resembles a bagging approach and might be analyzed in a
similar way. The objective function of the bipartite ranking problem
\cite{clemencon-2008-ranking-rates} (equivalently, AUC maximization)
is naturally framed as an optimization of a U-statistic. The ranking
problem discussed here is more general, bipartite ranking does not
consider the impact of a query context nor does it consider graded
relevance levels.

\subsection{Ranking and Empirical Minimization of $U$-Statistics}
Cl\`{e}men\c{c}on et al. \cite{clemencon-2008-ranking-rates} explore
fast rates for empirical risk minimization of $U$-statistics based
loss functions of order $2$. In particular, they show that under
suitable noise conditions, empirical risk minimization of the
bipartite ranking loss gives fast rates dependent on the exponent in
the satisfied noise condition. The low-noise condition is similar to
that defined elsewhere (see Boucheron et
al. \cite{boucheron-2005-fast-rates}) adapted for the $U$-statistics case:

\[\text{Var}(h_r(X, Y)) \leq c \Lambda(r)^\alpha\]

where $c$ is some positive constant and $\alpha \in [0, 1]$ is the coefficient determining the low-noise behavior. $h_r$ is the kernel defined in the Hoeffding decomposition of the empirical excess risk. Let $\Lambda(r)$ be the excess risk of a pairwise ranking function $r$:

\begin{equation}
  \label{eq:kernel}
  h_r(x,y) = \mathbb{E}(\mathbb{I}((y-Y)r(x, X) < 0) - \mathbb{I}((y-Y)r^*(x, X) < 0)) - \Lambda(r)
\end{equation}

\subsubsection{Main results}

Theorem 5 gives conditions under which we can achieve fast rates of convergence. The proof relies on 3 pieces:

\begin{itemize}
  \item The Hoeffding decomposition of the empirical excess risk into a sum of iid terms and a degenerate $U$-statistic of order $2$.
  \item A moment inequality for $U$-processes of order $2$ (Theorem 11) to show that the degenerate component of the decomposition is small with high probability.
  \item A bound on the excess risk in the iid case (Theorem 8.3 from Massart \cite{massart-2007-model-selection}).\footnote{It's still unclear to me how exactly Theorem 8.3 of Massart is applied. Is the moment bound used directly? In that case, how do we get $\delta$ in the statement of theorem 5 from $\delta/2$ in the proof? If the tail bound is used, then isnt Theorem 5 missing a constant term ($\kappa \epsilon^2_*$ in Massart)?}
\end{itemize}

Note that the moment inequality of Theorem 11 is largely based on the moment inequalities from Boucheron et al. \cite{boucheron-2005-moment-inequalities}. The proof is essentially application of the bounds present there in addition to a few applications of decoupling and un-decoupling steps (see de la Pe\~{n}a et al. chapter~3 \cite{de-la-pena-1999-decoupling}).

The bound in Theorem 5 is stated in terms of expectations of Rademacher averages and chaoses; the corollaries show some examples on how to bound these terms for the case of VC classes.


\textbf{Questions}

\begin{itemize}
  \item Can we extend the results of Clemencon et al.~\cite{clemencon-2008-ranking-rates} to listwise ranking with aggregation (as framed by Duchi)? The next step would be to apply a Hoeffding decomposition to the empirical risk \ref{eq:er} for \(k=3\) and see if the terms can be controlled similarly.
    \begin{itemize}
      \item First we must determine an analogous condition to Assumption 4 from \cite{clemencon-2008-ranking-rates} for the multi-partite case.
    \end{itemize}
  \item What is the next step in deriving fast rates for the case of bipartite ranking with a surrogate? Can Proposition 12 from Cl\`{e}men\c{c}on et al.~\cite{clemencon-2008-ranking-rates} be tightened?
    \begin{itemize}
      \item Cl\`{e}men\c{c}on et al. mention that Blanchard et al. \cite{blanchard-2003-rates} have explored this for binary classification.
    \end{itemize}
  \item How can we unify the notions of aggregation from Duchi and normalization from Ravikumar?
  \item Is the model selection problem more interesting? In particular for the case of listwise ranking in an online setting.
  \item From the conclusion of Duchi: investigate aggregation functions that yield non-point distributions as \(k\rightarrow \infty\). For example, scaling the Thurstone–Mosteller least-squares solutions by \(\sqrt k\) to achieve asymptotic normality could lead to more robust solutions.
  \item Ravikumar et al.~\cite{ravikumar-2011-ranking} stress the importance of normalizing the relevance labels by the maximum possible NDCG to ensure `robustness' to less likely but influential samples. Can this normalization be estimated with an EM type or other iterative approach?
\end{itemize}

\bibliography{summary}{}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
